---
title: "Adv Data Prog with R - Homework 1"
author: "Giyo Varghese"
date: "June 2, 2025"
format: pdf
---

# Introduction

The **`economics`** dataset contains monthly U.S. economic indicators from January 1967 to December 2015. In this report, we focus on:

- **`date`** – Month/year of each observation  
- **`unemploy`** – Number of unemployed (in thousands)  

---

# Dataset Description

The **`economics`** dataset includes:

- `date`: The date of each monthly observation (from 1967-01 to 2015-12).  
- `psavert`: Personal savings rate (percentage).  
- `pop`: Total U.S. population (in thousands).  
- `pce`: Personal consumption expenditures (in billions of dollars).  
- `unemploy`: Number of unemployed individuals (in thousands).  
- `uempmed`: Median duration of unemployment (in weeks).  

> *“Here, we use only **`date`** and **`unemploy`**, but knowing the other columns helps understand the broader context of the dataset.”*

---


```{r}
#| echo: false
#| warning: false
#| message: false
# Loading ggplot2
library(ggplot2)

# Importing 'economics' dataset
data(economics)
```

# Plotting

In this section, we create a **`line chart`** of U.S. monthly unemployment from January 1967 to December 2015

```{r}
#| echo: false
#| warning: false
#| message: false
# Create a line chart of 'unemploy' v/s 'date'
ggplot(economics, aes(x = date, y = unemploy)) +
  geom_line(color = "red", size = 0.8) + #First Change
  labs(
    title = "U.S. Unemployment",
    x = "Year",
    y = "Unemployed"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
    axis.title  = element_text(size = 12)
  )

```
Unemployment remained at low level at late 1960s before the spikes around in 1975(~8,100) and in 1982(~12,000), And after 1990s, a peak found near 8,500 and it fell until the 2001 downturn, then soared to 14,000 in 2008-09(maybe of recession). Post 2009, the unemployment steadily declined to around 8,200 by 2015.
